# Ass 1

========================================================================================================



This report will try to answer the questions of Assignment 2 of the course Reproducible Research
 

### Loading and preprocessing the data

To load the data i just used the code below. The dataset "dataperyear" has the aggregated data per date
```{r, echo=TRUE, results='hide'}
data<-read.csv("activity.csv")
dataperyear<- aggregate(data$steps,by=data["date"],FUN=sum, na.rm=TRUE)
```


### What is mean total number of steps taken per day?

**Make a histogram of the total number of steps taken each day**

```{r, echo=TRUE}
hist(dataperyear$x)

```



The code to calculate the mean and median total number of steps taken per day is the below. 

```{r, echo=TRUE}

mean<- aggregate(data$steps,by=data["date"],FUN=mean, na.rm=TRUE)

median<- aggregate(data$steps,by=data["date"],FUN=median, na.rm=TRUE)

```

The outcome was
1. Mean is ` r mean `
2. Median is ` r median `

### What is the average daily activity pattern?

**Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)**

```{r, echo=TRUE, results='hide'}
mean2<- aggregate(data$steps,by=data["interval"],FUN=mean, na.rm=TRUE)
```

```{r, echo=TRUE}
plot(mean2$interval, mean2$x, type="l", xlab= "interval", ylab= "Average number of steps taken", col="green" , lwd=2)
```


**Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**

```{r, echo=TRUE, results='hide'}
which.max(mean2$x)

s<-mean2[which.max(mean2$x),1]

```

The 5-minute interval which contains the max number of steps is ` r s `

### Imputing missing values

**Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)**

```{r, echo=TRUE}
library(plyr)
ddply(data,.(date), summarise, Count_Missing= sum(is.na(steps)), Count_No_Missing= sum(!is.na(steps)))
```

**Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.**

In this case i used the mean of 5-minute interval. Firstly i create a function which replace with the mean.
Then i used the function ddply. The code is the below

```{r, echo=TRUE, results='hide'}


impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))


#transform the data
data_trans<-ddply(data,.(interval), transform, steps= impute.mean(steps))

data_trans[order(data_trans$date), ] #plyr orders by group so we have to reorder

#onvert do date
data_trans$date<-as.Date(data_trans$date)


#create list with weekends

weekend<-c("Saturday", "Sunday")

#create one more column with weekdays

data_trans$weekday <- weekdays(data_trans$date, abbreviate=FALSE)

#create one more column with TRUE, FALSE regarding whether is weekend


data_trans$IsWeekend <- data_trans$weekday %in% weekend


```

The plot is the following

```{r, echo=TRUE}

#take the average per interval and per whether is weekend


mean3<- aggregate(data_trans$steps,by=list(data_trans$interval, data_trans$IsWeekend),FUN=mean, na.rm=TRUE)

#rename the column names
colnames(mean3)<-c("interval", "IsWeekend", "steps")
library(ggplot2)
p<-ggplot(mean3, aes(x=interval,y=steps))+geom_line(aes(color=IsWeekend),size=1.5)+facet_wrap(~IsWeekend, nrow=2)

print(p)

```